---
title: "Neushul Redux Notebook"
subtitle: "Pilot and Beyond"
author: "Ross Whippo"
date: 'Created: 2021-07-01  ; (Updated: `r Sys.Date()`)'
output:
  html_document:
    code_folding: hide
    df_print: kable
    fig_caption: yes
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float: true
  pdf_document:
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'

---



<style>
.column-left{
  float: left;
  width:48%;
  text-align: left;
}


.column-right{
  float: right;
  width: 48%;
  text-align: left;
}

.column-All{
  float: left;
  width:100%;
  text-align: left;
}
</style>




```{r setup, include=FALSE}

## Start with tidyverse library, and some default chunk options

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.align = 'center')


## use the following to manually set the working directory of RMarkdown documents. Note that this needs to be defined in this 'setup' r code chunk and will not work if placed elsewhere

# knitr::opts_knit$set(root.dir = '//60451-desktop/MRGIS/Biology/2021 Synthesis/Site_Summaries')

```

```{r, echo = FALSE}

## Load Libraries

library(tidyverse)
library(gt)
library(viridis)
library(ggmap)
library(algaeClassify)
library(vistime)
library(plotly)
library(rmarkdown)
library(DT)


theme_set(theme_classic())

```

```{r}
# Import all data and manage for analyses/visualizations



```

<!-- 

TODO: 


NOTES: 
Created in 2021 from KEEN_markdown_notebook.Rmd

-->

# General Notes

Funded in part by the Richard R. and Megumi F. Strathmann Endowed Fellowship at Friday Harbor Labs, and the PSA award I got.

# Field Notes

## General Field Notes

### 2021-07-01

Began data collection as part of the FHL Marine Botany final project. 

## Surveys

### Dropcam

2021-07-01

Finished about 40% of first dropcam survey from rowboat. Need to enter metadata into R workbook. Most of the footage was good, GoPro ran about an hour before the 32GB card was full. Need to make sure the light is adjusted properly between each deployment and that the float attached to the GoPro is clear of the lens. 

2021-07-07

Finished the remaining transects except for 27. Did it all from a power boat which was much easier. 

### Snorkel

2021-07-06

Collected snorkel UPC and video data, made about 1/4 of the way up the island from the southern point. Tide was fairly high so mostly captured fucus and ulva. Transect line was run ~1.5 m below surface.


### SCUBA

2021-06-29

Tested methods for data collection via SCUBA with GoPro, transect tape, and UPC datasheet. Need to: wear ankle weights, make longer cord for slate, make place on slate to hold pencil, make intervals along transect pre-written, have letter circle for substrate type, add abbreviations of algae at bottom of sheet for reference.

2021-07-08

First full SCUBA surey of site 33. Bathymetery was different than indicated in Neuhsul figure. Hit a wide flat area near the end of the transect before another large dropoff. Lots of walls. I think we may have drifted a bit further N than the original transect.

2021-07-09

Transects of 3 and 27 done. 3 went fairly smoothly. The transect line for 27 detached as we ran it so it displaced the start of the transect about 100ft from the shore. I was able to track the distance as we returned so I think I can correct the measurements on the data sheet. The shallows were very bad vis when we returned so didn't even attempt video there. 

3 - fairly steep slope at the begninning going into a gently plain. Sheer wall near the interface of the two. Agarum deeper was attached to cobbles, maybe blown around?
27 - Soft sediment and lots of plants attached to cobble, shell, and small rocks. Did see at least two juvenile Nereo. Primarily gaudichaudii, costaria, desmarestia, patchy ulva, and sargassum in the shallows. Very long, low-sloping soft sediment plain. Almost no Zostera (only detrital). 

# Analysis Notes

## CoralNet

2021-07-05

Created labelset for photos and cropped annotation area to exlude most of the frame. Began identifying easy algae.

2021-07-07

Finished ID of first 17 photos. Calling sediment-covered rock 'bare rock'. 

# Acknowledgements

Snorkel: Gabi, Miranda
Training: Pema
Boat Tending: Kristy, Izzy
Dive Buddy: Kindall
Money: Tom and Wilson, PSA, FHL





```{r, echo = FALSE}

## Load Libraries


```



```{r}
# Import all data and manage for analyses/visualizations




```




<!-- end of document -->
